---
#title: "notebook_gene"
output:
  cleanrmd::html_document_clean:
    theme: almond
    css: css/headings.css
    lib_dir: R
---

```{r, include=FALSE}
#Librerias a utilizar
library(tidyverse)
library(ghql)
library(jsonlite)
library(dplyr)
library(httr)
library(tidyr)
library(ggplot2)
library(DT)
library(knitr)
library(htmltools)
library(htmlwidgets)
library(rmarkdown)
library(webshot2)
```

```{r setup, include=FALSE}
# To obtain the list of .r files in the "R" folder
archivos <- list.files("R", pattern = "\\.R$", full.names = TRUE)
# To load the functions from each file
lapply(archivos, source)

# To obtain the list of .png files in the "images" folder
images <- list.files("images", pattern = "\\.png$", full.names = TRUE)

# To obtain the list of .csv files in the "datas" folder
datas <- list.files("datas", pattern = "\\.csv$", full.names = TRUE)

```
```{r, include=FALSE}
regulondb_conn <- connect_database()

get_gene_general_statistics <- get_gene_general_statistics()
capture_table_data(get_gene_general_statistics, "get_gene_general_statistics")
capture_table_png(get_gene_general_statistics,"get_gene_general_statistics")

get_calculate_gene_intervals <- get_calculate_gene_intervals()
capture_table_data(get_calculate_gene_intervals, "get_calculate_gene_intervals")
capture_table_png(get_calculate_gene_intervals, "get_calculate_gene_intervals")

get_tfs_regulations_genes <- get_tfs_regulations_genes()
capture_table_data(get_tfs_regulations_genes,"get_tfs_regulations_genes")
capture_table_png(get_tfs_regulations_genes,"get_tfs_regulations_genes")

```

<script>
function download_resources(fileUrl) {
  var link = document.createElement('a');
  link.href = fileUrl;

  // Extraer el nombre del archivo de la URL para establecerlo como nombre de descarga
  var fileName = fileUrl.split('/').pop();
  link.download = fileName;
  
  document.body.appendChild(link);
  link.click();
  document.body.removeChild(link);
}
</script>


### Gene Overview Statistics

DESCRIPTION

This gene overview displays some general data and distributions related to the genes stored in RegulonDB.

<a onclick="download_resources('images/get_gene_general_statistics.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Image</a>
   
<a onclick="download_resources('datas/get_gene_general_statistics.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_gene_general_statistics, rownames = FALSE, 
          options = list(dom = 't', ordering = FALSE, lengthMenu = list(c(-1))))
#knitr::kable(get_gene_general_statistics)
```

### Gene Size

This section elucidates the spectrum of gene sizes in *E.coli*. The sizes, calculated for each individual gene, are grouped into specific size ranges, which are detailed in the accompanying table. To provide a visual representation of these size ranges and their prevalence, a graph displaying the frequency of each range is also included.


<a onclick="download_resources('images/get_calculate_gene_intervals.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Image</a>
   
<a onclick="download_resources('datas/get_calculate_gene_intervals.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_calculate_gene_intervals)
ggplot(get_calculate_gene_intervals, aes(x = `Size of Genes`, y = Total)) +
  geom_bar(stat = "identity", fill = "#585554") +
  geom_text(aes(label = `Total`, y = Total + 10), vjust = 0.5 , size = 3, angle = 90) +
  labs(title = "Distribution of the gene sizes.") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### Distribution of the Number of Transcription Factors Regulating Genes

At RegulonDB, we archive detailed information on gene regulation. Transcription factors (TFs) bind to the promoter region, thereby regulating the transcriptional unit, which comprises a promoter with binding sites, a set of transcribed genes, and a terminator region where the transcription process concludes. This section provides a count of how many regulators or TFs are involved in the transcription of the genes.

<a onclick="download_resources('images/get_tfs_regulations_genes.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Image</a>
   
<a onclick="download_resources('datas/get_tfs_regulations_genes.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_tfs_regulations_genes)
ggplot(get_tfs_regulations_genes, aes(x = `Total of TFs that regulate the gene`, y = `Total of Genes`)) +
  geom_bar(stat = "identity", fill = "#585554") +
  geom_text(aes(label = `Total of Genes`), vjust = -0.5 , size = 3) +
  labs(title = "Number of TF regulating the genes.") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```


