
---
title: "Notebook Transcription Factors"
output: 
  html_document
    

---
```{r, include = FALSE}
library(regutools)
library(ghql)
library(jsonlite)
library(dplyr)
library(httr)
library(tidyr)
library(ggplot2)
library(DT)
library(knitr)
library(htmltools)
library(htmlwidgets)
library(shinyjs)
library(rmarkdown)
library(data.table)
library(cleanrmd)

```

```{r, include = FALSE}
###connection
regulondb_conn <- connect_database()

###genes by tf
get_genes_by_tf <- get_genes_by_tf()
get_general_frequency_genes_by_tf <- get_general_frequency(get_genes_by_tf,data.getAllRegulon.data.transcriptionFactor$name, cantidad,n_tf)
generate_graphic_points_genes_by_tfs <- generate_graphic_points(get_general_frequency_genes_by_tf,	n_tf, cantidad, "Number regulated genes ", "Number of tf","Number of genes regulated by tf")

###tus by tf

get_tus_by_tfs <- get_tus_by_tfs()
get_general_frequency_tus_by_tfs <- get_general_frequency(get_tus_by_tfs,name, cantidad,n_tus)
 generate_graphic_points_tus_by_tfs <- generate_graphic_points(get_general_frequency_tus_by_tfs,cantidad,n_tus ,"numbers tf ", "numbers tus","Number of tus regulated by tf ")

```


## Introduction
##### **introduccion proporcionada por el curador**

## Examples
**pendientes a  la bd**
** imagenes o links para cada ejemplo proporcionado por el curador**
#### Holo Suppressor
#### apo repressor
#### Holo trigger
#### apo activator

##  transcription factor

#### **How many genes are regulated by each TF? / No of genes regulated by TF (graph)**
```{r, warning = FALSE, echo = FALSE}

get_genes_by_tf
get_general_frequency_genes_by_tf
generate_graphic_points_genes_by_tfs
```

###### **pie de grafica proporcionada por el curador** 


####  **How many genes are regulated by each TF organized by family? / No of TF-regulated genes grouped by family (graph)**

####  **How many are regulated by each TF? / No of Tus regulated by TF (graph)**
```{r, warning = FALSE, echo = FALSE}
get_tus_by_tfs
get_general_frequency_tus_by_tfs 
generate_graphic_points_tus_by_tfs
```

###### pie de grafica proporcionada por el curador** 


####  **What is the number of tfrs per tf ?(graphs)**

####  **What is the size distribution of the TFRs? (Graph )**

####  **Distribution of distances from the center of the TFBSs to the promoter and to the beginning of the first gene: activators, and repressors (only for RIs tf-promoter)**

####  **Distribution of distances of the TFBSs to the gene: activators, and repressors (tf-promoter, tf-tu, tf-gene)**

####  **Distribution of distances of TFBSs to the gene by TF: activators, and repressors**

####  **What is the number of RIs of each type: TF-promoter, TF-TU, TF-gene and their confidence level? (Graph and Table)**

####  **How is the distribution No of Regulators and RIs by TU/promoter/genes (activators, repressors)?(Graphic)**
