---
#title: "notebook_gene"
output:
  cleanrmd::html_document_clean:
    theme: almond
    css: css/headings.css
    lib_dir: R
---

```{r, include=FALSE}
#Librerias a utilizar
library(tidyverse)
library(ghql)
library(jsonlite)
library(dplyr)
library(httr)
library(tidyr)
library(ggplot2)
library(DT)
library(knitr)
library(htmltools)
library(htmlwidgets)
library(rmarkdown)
library(webshot2)
```

```{r setup, include=FALSE}
# To obtain the list of .r files in the "R" folder
archivos <- list.files("R", pattern = "\\.R$", full.names = TRUE)
# To load the functions from each file
lapply(archivos, source)

# To obtain the list of .png files in the "images" folder
images <- list.files("images", pattern = "\\.png$", full.names = TRUE)

# To obtain the list of .csv files in the "datas" folder
datas <- list.files("datas", pattern = "\\.txt$", full.names = TRUE)

# To obtain the list of .png files in the "images" folder
graphics <- list.files("graphics", pattern = "\\.png$", full.names = TRUE)
```

```{r, include=FALSE}
regulondb_conn <- connect_database()

# 1
get_gene_general_statistics <- get_gene_general_statistics()
capture_table_data(get_gene_general_statistics, "data_get_gene_general_statistics")
capture_table_png(get_gene_general_statistics,"image_get_gene_general_statistics")

# 2
get_calculate_gene_intervals <- get_calculate_gene_intervals()
capture_table_data(get_calculate_gene_intervals, "data_get_calculate_gene_intervals")
capture_table_png(get_calculate_gene_intervals, "image_get_calculate_gene_intervals")

# 3
get_tfs_regulations_genes <- get_tfs_regulations_genes()
capture_table_data(get_tfs_regulations_genes,"data_get_tfs_regulations_genes")
capture_table_png(get_tfs_regulations_genes,"image_get_tfs_regulations_genes")

# 4
get_promoters_regulations_genes <- get_promoters_regulations_genes()
capture_table_data(get_promoters_regulations_genes,"data_get_promoters_regulations_genes")
capture_table_png(get_promoters_regulations_genes,"image_get_promoters_regulations_genes")

get_tus_regulations_genes <- get_tus_regulations_genes()
capture_table_data(get_tus_regulations_genes,"data_get_tus_regulations_genes")
capture_table_png(get_tus_regulations_genes,"image_get_tus_regulations_genes")

dfs_promoters_tus_genes_joined <- dfs_promoters_tus_genes_joined(
  get_promoters_regulations_genes,"Promoters",get_tus_regulations_genes,"TUs")

# 5
get_sigmafactor_regulations_genes <- get_sigmafactor_regulations_genes()
capture_table_data(get_sigmafactor_regulations_genes,"data_get_sigmafactor_regulations_genes")
capture_table_png(get_sigmafactor_regulations_genes,"image_get_sigmafactor_regulations_genes")

# 6
get_class_molecularfunction_by_genes <- get_class_molecularfunction_by_genes()
capture_table_data(get_class_molecularfunction_by_genes,"data_get_class_molecularfunction_by_genes")

get_class_cellularcomponent_by_genes <- get_class_cellularcomponent_by_genes()
capture_table_data(get_class_cellularcomponent_by_genes,"data_get_class_cellularcomponent_by_genes")

get_class_biologicalprocess_by_genes <- get_class_biologicalprocess_by_genes()
capture_table_data(get_class_biologicalprocess_by_genes,"data_get_class_biologicalprocess_by_genes")

```

```{=html}
<script>
function download_resources(fileUrl) {
  var link = document.createElement('a');
  link.href = fileUrl;

  // Extraer el nombre del archivo de la URL para establecerlo como nombre de descarga
  var fileName = fileUrl.split('/').pop();
  link.download = fileName;
  
  document.body.appendChild(link);
  link.click();
  document.body.removeChild(link);
}
</script>
```
### Gene Overview Statistics

DESCRIPTION

This gene overview displays some general data and distributions related to the genes stored in RegulonDB.

<a onclick="download_resources('images/image_get_gene_general_statistics.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Image</a>

<a onclick="download_resources('datas/data_get_gene_general_statistics.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_gene_general_statistics, rownames = FALSE, 
          options = list(dom = 't', ordering = FALSE, lengthMenu = list(c(-1))))
#knitr::kable(get_gene_general_statistics)
```

### Gene Size

This section elucidates the spectrum of gene sizes in *E.coli*. The sizes, calculated for each individual gene, are grouped into specific size ranges, which are detailed in the accompanying table. To provide a visual representation of these size ranges and their prevalence, a graph displaying the frequency of each range is also included.

<a onclick="download_resources('images/image_get_calculate_gene_intervals.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Image</a>

<a onclick="download_resources('datas/data_get_calculate_gene_intervals.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_calculate_gene_intervals, rownames = FALSE)
```

<a onclick="download_resources('graphics/graphic_get_calculate_gene_intervals.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Graphic</a>

```{r, warning=FALSE, echo=FALSE}
graphic_get_calculate_gene_intervals <- ggplot(get_calculate_gene_intervals, 
                                               aes(x = `Size of Genes`, y = Total)) +
  geom_bar(stat = "identity", fill = "#585554") +
  geom_text(aes(label = `Total`, y = Total + 10), vjust = 0.5 , size = 3, angle = 90) +
  labs(title = "Distribution of the gene sizes.") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

graphic_get_calculate_gene_intervals

capture_graphic_png(graphic_get_calculate_gene_intervals,"graphic_get_calculate_gene_intervals")
```

### Distribution of the Number of Transcription Factors Regulating Genes

At RegulonDB, we archive detailed information on gene regulation. Transcription factors (TFs) bind to the promoter region, thereby regulating the transcriptional unit, which comprises a promoter with binding sites, a set of transcribed genes, and a terminator region where the transcription process concludes. This section provides a count of how many regulators or TFs are involved in the transcription of the genes.

<a onclick="download_resources('images/image_get_tfs_regulations_genes.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Image</a>

<a onclick="download_resources('datas/data_get_tfs_regulations_genes.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_tfs_regulations_genes, rownames = FALSE)
```

<a onclick="download_resources('graphics/graphic_get_tfs_regulations_genes.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Graphic</a>

```{r, warning=FALSE, echo=FALSE}
graphic_get_tfs_regulations_genes <- ggplot(get_tfs_regulations_genes, 
                                            aes(x = `Total of TFs that regulate the gene`, 
                                                y = `Total of Genes`)) +
  geom_bar(stat = "identity", fill = "#585554") +
  geom_text(aes(label = `Total of Genes`), vjust = -0.5 , size = 3) +
  labs(title = "Number of TFs regulating the genes.") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

graphic_get_tfs_regulations_genes

capture_graphic_png(graphic_get_tfs_regulations_genes,"graphic_get_tfs_regulations_genes")
```

### Distribution of the Number of TUs and Promoters by Gene

DESCRIPTION \*\*

#### Distribution of Total promoters per gene

<a onclick="download_resources('images/image_get_promoters_regulations_genes.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Image</a>

<a onclick="download_resources('datas/data_get_promoters_regulations_genes.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_promoters_regulations_genes, rownames = FALSE, 
          options = list(dom = 't', ordering = FALSE, lengthMenu = list(c(-1)),
                         columnDefs = list( list(targets = "_all", className = "dt-center"))))
```

#### Distribution of Total Transcription Units (TUs) per gene.

<a onclick="download_resources('images/image_get_tus_regulations_genes.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Image</a>

<a onclick="download_resources('datas/data_get_tus_regulations_genes.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_tus_regulations_genes, rownames = FALSE, 
          options = list(dom = 't', ordering = FALSE, lengthMenu = list(c(-1)),
                         columnDefs = list( list(targets = "_all", className = "dt-center"))))
```

<a onclick="download_resources('graphics/graphic_dfs_promoters_tus_genes_joined.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Graphic</a>

```{r, warning=FALSE, echo=FALSE}
graphic_dfs_promoters_tus_genes_joined <- ggplot(dfs_promoters_tus_genes_joined, 
                                                 aes(x = Number, 
                                                     y = `Total of Genes`, 
                                                     color = factor(Type))) +
  ggtitle( "Distribution of Genes Regulated by Number of Promoters and TUs.") +
  xlab("Total Number of Promoters and TUs") +
  ylab("Total Number of Genes") +
  geom_line() +
  scale_color_manual(values = c("Promoters" = "blue", "TUs" = "red"), name = "Type") +
  scale_x_continuous(breaks = seq(0, max(dfs_promoters_tus_genes_joined$Number), by = 1)) +
  theme_minimal()+
  theme(plot.background = element_rect(fill = "white", color = "white"))

graphic_dfs_promoters_tus_genes_joined

capture_graphic_png(graphic_dfs_promoters_tus_genes_joined,"graphic_dfs_promoters_tus_genes_joined")
```

### Distribution of Sigma Factor Regulatory Genes

DESCRIPTION **

<a onclick="download_resources('images/image_get_sigmafactor_regulations_genes.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Image</a>

<a onclick="download_resources('datas/data_get_sigmafactor_regulations_genes.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_sigmafactor_regulations_genes, rownames = FALSE,
          options = list(dom = 't', ordering = FALSE, lengthMenu = list(c(-1)),
                         columnDefs = list( list(targets = "_all", className = "dt-center"))))
```

<a onclick="download_resources('graphics/graphic_get_sigmafactor_regulations_genes.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Graphic</a>

```{r, warning=FALSE, echo=FALSE}
graphic_get_sigmafactor_regulations_genes <- ggplot(get_sigmafactor_regulations_genes, 
                                                    aes(x = `Total of sigma factors regulating genes`, 
                                                        y = `Total of Genes`)) +
  geom_bar(stat = "identity", fill = "#585554") +
  geom_text(aes(label = `Total of Genes`), vjust = -0.5 , size = 3) +
  labs(title = "Total of Sigma Factors regulating the genes.") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

graphic_get_sigmafactor_regulations_genes

capture_graphic_png(graphic_get_sigmafactor_regulations_genes,"graphic_get_sigmafactor_regulations_genes")
```


### Distribution of the Number of Functional Classes by Gene

DESCRIPTION \*\*

#### Table of Number of Molecular Function Classes by Gene

<a onclick="download_resources('datas/data_get_class_molecularfunction_by_genes.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_class_molecularfunction_by_genes, rownames = FALSE)
```

<a onclick="download_resources('graphics/graphic_get_class_molecularfunction_by_genes.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Graphic</a>

```{r, warning=FALSE, echo=FALSE}
graphic_get_class_molecularfunction_by_genes <- ggplot(get_class_molecularfunction_by_genes, 
                                                       aes(x = `Class Molecular Function`, 
                                                           y = `Total of Genes`)) +
  geom_bar(stat = "identity", fill = "#585554") +
  labs(title = "Distribution of Molecular Function Class.") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_blank())

graphic_get_class_molecularfunction_by_genes

capture_graphic_png(graphic_get_class_molecularfunction_by_genes,"graphic_get_class_molecularfunction_by_genes")
```

#### Table of Number of Cellular Component Classes by Gene

<a onclick="download_resources('datas/data_get_class_cellularcomponent_by_genes.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_class_cellularcomponent_by_genes, rownames = FALSE)
```

<a onclick="download_resources('graphics/graphic_get_class_cellularcomponent_by_genes.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Graphic</a>

```{r, warning=FALSE, echo=FALSE}
graphic_get_class_cellularcomponent_by_genes <- ggplot(get_class_cellularcomponent_by_genes, 
                                                       aes(x = `Class Cellular Component`, 
                                                           y = `Total of Genes`)) +
  geom_bar(stat = "identity", fill = "#585554") + 
  labs(title = "Distribution of Cellular Component Class.") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_blank())

graphic_get_class_cellularcomponent_by_genes

capture_graphic_png(graphic_get_class_cellularcomponent_by_genes,"graphic_get_class_cellularcomponent_by_genes")
```

#### Table of Number of Biological Processes by Gene

<a onclick="download_resources('datas/data_get_class_biologicalprocess_by_genes.txt')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Data</a>

```{r, warning=FALSE, echo=FALSE}
datatable(get_class_biologicalprocess_by_genes, rownames = FALSE)
```

<a onclick="download_resources('graphics/graphic_get_class_biologicalprocess_by_genes.png')" 
   style="font-style: italic; text-decoration: none; cursor: pointer;">Download Graphic</a>

```{r, warning=FALSE, echo=FALSE}
graphic_get_class_biologicalprocess_by_genes <- ggplot(get_class_biologicalprocess_by_genes, 
                                                       aes(x = `Class Biological Process`, 
                                                           y = `Total of Genes`)) +
  geom_bar(stat = "identity", fill = "#585554") +
  labs(title = "Distribution of Biological Process.") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_blank())

graphic_get_class_biologicalprocess_by_genes

capture_graphic_png(graphic_get_class_biologicalprocess_by_genes,"graphic_get_class_biologicalprocess_by_genes")
```
